\textsf{Are large language models superhuman chemists?}

### 2024-03-30 src/tex/appendix.tex
remove scaling plot and instead just add numbers in the text

To further investigate the performance of the models, we also compared the performance on different data sources.
Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some of the semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.

@@

\subsection{Performance as a function of molecular features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}). 
The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.

@@

\subsection{Influence of scale}
To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}

\Cref{fig:llama_performance} shows the performance of the LLaMA (chat) models with different parameter counts on the \chembench corpus.

\begin{figure}[htb]
    \centering 
    \includegraphics{figures/llama_performance.pdf}
    \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
    \label{fig:llama_performance}
    \script{analyze_llama_performance.py}
\end{figure}

@@

\subsection{Leaderboard}
Our leaderboard is based on the toolchain developed for Matbench.\autocite{Dunn_2020} 
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

>>>>

To further investigate the performance of the models, we also compared the performance on different data sources.
Compared to topics, this is a more fine-grained analysis, as topics can be composed of questions from different sources.
In \Cref{fig:performance_per_topic}, we see that the performance of the models varies significantly between the different data sources.
Interestingly, the performance of the models on questions sourced based on textbooks seems to be better for our models than some semi-programmatically created tasks, such as questions about the number of signals in an \gls{nmr} spectrum.

@@

\subsection{Performance as a function of molecular features}
To better understand if the performance of the models is correlated with specific features of the molecules, we analyzed the performance of the models as a function of the number of atoms and the complexity of the molecules.
\Cref{fig:correlation_plot_is_number_nmr_peaks_complexity} shows that the performance of the models is not correlated with the complexity of the molecules but rather with the number of atoms (\Cref{fig:correlation_plot_is_number_nmr_peaks_num_atoms}, similar trivial correlation for \Cref{fig:correlation_plot_is_electron_counts_num_atoms}). 
The corresponding Spearman correlation coefficients are listed in \Cref{tab:correlation_coefficients}.

@@

\subsection{Influence of model scale}
To obtain first insights in how the performance \glspl{llm} depends on scale, we tested the \glspl{llm} of the LLaMA series. 
Interestingly, we find that the 7B and 70B models perform comparably, with the 7B showing lower performance ().

Note that such analyses are difficult as models are typically not directly comparable in terms of dataset and training protocol.\autocite{biderman2023pythia}

@@

\subsection{Leaderboard}
Our leaderboard is based on the tool chain developed for Matbench.\autocite{Dunn_2020} 
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

### 2024-03-30 src/tex/ms.tex
remove scaling plot and instead just add numbers in the text

We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see XX).
This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}

>>>>

We also gain insight into the models' struggles with chemical reasoning tasks when we examine their performance as a function of molecular descriptors.
If the model would answer questions based on reasoning about the structures, one would expect the performance to depend on the complexity of the molecules.
However, we find that the models' performance does not correlate with complexity indicators but rather trivially with the size of the compounds (see \Cref{tab:correlation_coefficients}).
This indicates that the models may not be able to reason about the structures of the molecules (in the way one might expect) but instead rely on retrieval of fragments of the training data.
Those observations mirror recent findings that \glspl{llm} struggle to \enquote{reverse} facts they have seen in training (i.e., generalize from \enquote{A has feature B} to \enquote{B is a feature of A}).\autocite{berglund2023reversal, zhu2023physics, allen2023physics, golovneva2024reverse}
